### 2024-03-27 doc/sections/mav-semantics.tex
[ tex ] fill in some bits

\begin{definition}
  \bob{residuation}
\end{definition}

The definition of $*$-autonomous category is due to Barr
\cite{Barr_1979}. For our purposes, we need the partial order
analogue, also called a CL algebra by Troelstra
\cite{Troelstra92:lll}.

@@

\begin{remark}
  If $\bullet$ is a join, or $\lhd$ is a meet, then all the conditions
  for a duoidal relationship are automatically met for any other
  monoid. \bob{fwd ref to where we use this.}
\end{remark}

@@

\bob{Change $;$ to $\lhd$}

\begin{definition}\label{defn:mav-algebra}
  An \emph{MAV-algebra} is a structure
  $(A, \leq, \otimes, ;, I, \lnot)$ with the following properties:
  \begin{enumerate}
  \item $(A, \leq, \otimes, I, \lnot)$ is $*$-autonomous and satisfies \emph{mix}.
  \item $(A, \leq, ;, I)$ is a pomonoid.
  \item $;$ is self dual: $\lnot (x ; y) = (\lnot x); (\lnot y)$.
  \item $(\otimes, I)$ is duoidal over $(;, I)$.
  \item $(A, \leq)$ has binary meets, which we write as $x \with y$.
  \end{enumerate}
\end{definition}

\begin{proposition}\label{prop:mav-algebra-consequences}
  Let $(A, \leq, \otimes, ;, I, \lnot)$ be a MAV-algebra.
  \begin{enumerate}
    \item There is another pomonoid structure $(\parr, I)$ on
          $(A, \leq)$, defined as
          $x \parr y = \lnot(\lnot x \otimes \lnot y)$.
    \item $(\otimes, I)$ and $(\parr, I)$ are linearly distributive
          \cite{cockett}:
          $x \otimes (y \parr z) \leq (x \otimes y) \parr z$.
    \item $(A, \leq)$ has binary joins, given by
          $x \oplus y = \lnot (\lnot x \with \lnot y)$.
    \item $\oplus$ distributes over $\otimes$:
          $x \otimes (y \oplus z) = ((x \otimes y) \oplus (x \otimes z)$.
    \item $\with$ distributes over $\parr$:
          $(x \parr z) \with (y \parr z) = (x \with y) \parr z$.
    \item $;$ is entropic over $\parr$:
          $(w \parr x); (y \parr z) \leq (w;y) \parr (x;z)$.
    \item $;$ is entropic over $\with$:
          $(w \with x); (y \with z) \leq (w ; y) \with (x; z)$.
    \item $\oplus$ is entropic over $;$:
          $(w ; x) \oplus (y ; z) \leq (x \oplus y) ; (x \oplus z)$.
  \end{enumerate}
\end{proposition}

@@

\begin{definition}\label{defn:mav-interpretation}
  Let $\mathit{At}$ be a set of atomic propositions. Given an
  MAV-algebra $(A, \leq, \otimes, \lhd, I, \lnot)$ and valuation
  $V : \mathit{At} \to A$, define the interpretation of MAV Formulas
  as follows: $\sem{\vUnit} = I$, $\sem{\alpha} = V(\alpha)$,
  $\sem{\vDual \alpha} = \lnot V(\alpha)$,
  $\sem{\vP\vTens\vQ} = \sem{\vP} \otimes \sem{\vQ}$,
  $\sem{\vP\vParr\vQ} = \sem{\vP} \parr \sem{\vQ}$,
  $\sem{\vP\vSeq\vQ} = \sem{\vP};\sem{\vQ}$,
  $\sem{\vP\vWith\vQ} = \sem{\vP} \with \sem{\vQ}$, and
  $\sem{\vP\vPlus\vQ} = \sem{\vP} \oplus \sem{\vQ}$.
\end{definition}

@@

We now turn to the self-dual duoidal structure required to interpret
the $\vSeq$ connective. First we transfer pomonoids from $(A, \leq)$
to self-dual pomonoids on $(\Chu(A, k), \sqsubseteq)$ provided they
interact well with $k$:
\begin{proposition}\label{prop:chu-self-dual}
  Let $(\lhd, j)$ be a pomonoid on $(A, \leq)$ such that $(\bullet, i)$ is entropic over 
  $k \lhd k \leq k$ and $j \leq k$, then
  $x ; y = (x^+ \lhd y^+, x^- \lhd y^-)$ and $J = (j, j)$ form a
  self-dual pomonoid on $\Chu(A, k)$.
\end{proposition}

\begin{proof}
  $x;y$ is well-defined because
  $(x^+ \lhd y^+) \bullet (x^- \lhd y^-) \leq (x^+ \bullet x^-) \lhd
  (y^+ \bullet y^-) \leq k \lhd k \leq k$. $J$ is well defined because
  $j \bullet j \leq j \leq k$. The pomonoid laws all transfer directly.
\end{proof}

@@

\begin{proposition}
  If $(\bullet, i)$ is entropic over $(\lhd, j)$ on $(A, \leq)$, and
  $(\lhd, j)$ satisfies the conditions of Proposition
  \ref{prop:chu-self-dual}, then $(\otimes, I)$ and $(;, J)$ are in a
  duoidal relationship on $\Chu(A, k)$.
\end{proposition}

>>>>

\begin{definition}
  A pomonoid $(\bullet, i)$ on a poset $(A, \leq)$ is \emph{left
    residuated} if there is a function
  $\leftblackspoon : A \times A \to A$ such that $x \bullet y \leq z$
  iff $x \leq z \leftblackspoon y$. The same structure is \emph{right
    residuated} if there is a function
  $\rightblackspoon : A \times A \to A$ such that $x \bullet y \leq z$
  iff $y \leq x \rightblackspoon z$. When $(\bullet, i)$ is
  commutative, the two definitions coincide, and we just say that the
  pomonoid is residuated.
\end{definition}
\bob{Check that the residuation definitions are the right way round}

The definition of $*$-autonomous category is due to Barr
\cite{Barr_1979}, and predates the introduction of Linear Logic. For
our purposes, we need the partial order analogue, also called a CL
algebra by Troelstra \cite{Troelstra92:lll}.

@@

\begin{remark}
  If $\bullet$ is a join, or $\lhd$ is a meet, then all the conditions
  for a duoidal relationship are automatically met.
\end{remark}

@@

We define MAV-algebras as the algebraic semantics of MAV. The
definition is a direct translation of the rules of MAV into
order-theoretic language, using the definitions we have seen so far.

\begin{definition}\label{defn:mav-algebra}
  An \emph{MAV-algebra} is a structure
  $(A, \leq, \otimes, \lhd, I, \lnot)$ with the following properties:
  \begin{enumerate}
  \item $(A, \leq, \otimes, I, \lnot)$ is $*$-autonomous and satisfies \emph{mix}.
  \item $(A, \leq, \lhd, I)$ is a pomonoid.
  \item $\lhd$ is self dual: $\lnot (x \lhd y) = (\lnot x)\lhd (\lnot y)$.
  \item $(\otimes, I)$ is duoidal over $(\lhd, I)$.
  \item $(A, \leq)$ has binary meets, which we write as $x \with y$.
  \end{enumerate}
\end{definition}

\begin{proposition}\label{prop:mav-algebra-consequences}
  Let $(A, \leq, \otimes, \lhd, I, \lnot)$ be a MAV-algebra.
  \begin{enumerate}
    \item There is another pomonoid structure $(\parr, I)$ on
          $(A, \leq)$, defined as
          $x \parr y = \lnot(\lnot x \otimes \lnot y)$.
    \item $(\otimes, I)$ and $(\parr, I)$ are linearly distributive
          \cite{cockett}:
          $x \otimes (y \parr z) \leq (x \otimes y) \parr z$.
    \item $(A, \leq)$ has binary joins, given by
          $x \oplus y = \lnot (\lnot x \with \lnot y)$.
    \item $\oplus$ distributes over $\otimes$:
          $x \otimes (y \oplus z) = ((x \otimes y) \oplus (x \otimes z)$.
    \item $\with$ distributes over $\parr$:
          $(x \parr z) \with (y \parr z) = (x \with y) \parr z$.
    \item $\lhd$ is entropic over $\parr$:
          $(w \parr x)\lhd (y \parr z) \leq (w\lhd y) \parr (x\lhd z)$.
    \item $\lhd$ is entropic over $\with$:
          $(w \with x)\lhd (y \with z) \leq (w \lhd y) \with (x\lhd z)$.
    \item $\oplus$ is entropic over $\lhd$:
          $(w \lhd x) \oplus (y \lhd z) \leq (x \oplus y) \lhd (x \oplus z)$.
  \end{enumerate}
\end{proposition}

@@

\begin{definition}\label{defn:mav-interpretation}
  Let $\mathit{At}$ be a set of atomic propositions. Given an
  MAV-algebra $(A, \leq, \otimes, \lhd, I, \lnot)$ and valuation
  $V : \mathit{At} \to A$, define the interpretation of MAV Formulas
  as follows: $\sem{\vUnit} = I$, $\sem{\alpha} = V(\alpha)$,
  $\sem{\vDual \alpha} = \lnot V(\alpha)$,
  $\sem{\vP\vTens\vQ} = \sem{\vP} \otimes \sem{\vQ}$,
  $\sem{\vP\vParr\vQ} = \sem{\vP} \parr \sem{\vQ}$,
  $\sem{\vP\vSeq\vQ} = \sem{\vP}\lhd\sem{\vQ}$,
  $\sem{\vP\vWith\vQ} = \sem{\vP} \with \sem{\vQ}$, and
  $\sem{\vP\vPlus\vQ} = \sem{\vP} \oplus \sem{\vQ}$.
\end{definition}

@@

We now turn to the self-dual duoidal structure required to interpret
the $\vSeq$ connective. First we transfer pomonoids from $(A, \leq)$
to self-dual pomonoids on $(\Chu(A, k), \sqsubseteq)$ provided they
interact well with $k$:
\begin{proposition}\label{prop:chu-self-dual}
  Let $(\lhd, j)$ be a pomonoid on $(A, \leq)$ such that $(\bullet, i)$ is entropic over 
  $k \lhd k \leq k$ and $j \leq k$, then
  $x \lhd y = (x^+ \lhd y^+, x^- \lhd y^-)$ and $J = (j, j)$ form a
  self-dual pomonoid on $\Chu(A, k)$.
\end{proposition}

\begin{proof}
  $x\lhd y$ is well-defined because
  $(x^+ \lhd y^+) \bullet (x^- \lhd y^-) \leq (x^+ \bullet x^-) \lhd
  (y^+ \bullet y^-) \leq k \lhd k \leq k$. $J$ is well defined because
  $j \bullet j \leq j \leq k$. The pomonoid laws all transfer directly.
\end{proof}

@@

\begin{proposition}
  If $(\bullet, i)$ is entropic over $(\lhd, j)$ on $(A, \leq)$, and
  $(\lhd, j)$ satisfies the conditions of Proposition
  \ref{prop:chu-self-dual}, then $(\otimes, I)$ and $(\lhd, J)$ are in a
  duoidal relationship on $\Chu(A, k)$.
\end{proposition}
